---
title: "Electric Vehicle Overview"
author: "Wenwei Fei"
date: 2021-09-10
categories: ["R"]
tags: ["ggplot2", "Vehicle Price Analysis","ggiraph"]
draft: FALSE
slug: "ev-analysis"
thumbnail: "images/birds/bird_5.jpg"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
```

```{r}
library(tidyverse)
library(ggiraph)
library(data.table)
library(scales)
library(ggpubr)
```

```{r}
filepath <- "D:/Projects/EVE/Eve/"
df_ev <- fread(paste0(filepath, "df_EV_clean.csv"))
df_ev[,.N,make]
df_ev <- df_ev[!make == "Mitsubishi"]
df_ev[, .N, body_type]
df_ev[, range_km := ifelse(str_detect(make, "Jaguar"), 446, range_km)]
df_ev[, my := as.factor(my)]
```

# Overview

Hello again! After scraping and clean, I can finally present this overview of pure electric vehicle analysis. 

You can find my work on my [Github](https://github.com/wwfeitim/Eve) and are welcome to study the data I collected. I converted the datasheet to a CSV file and named it "df_EV_clean.csv". 

The data include 881 pure electric cars on sale in Aus Market with variables including price, model years, makes, body types, battery power, travel range and odometers. 

I hope you can find some answers from this post:

- What brands provide pure EVs in the Australian market?
- How many budgets do I need to buy an EV?
- What features and metrics of an EV are important? 
- How do battery power and travel range affect the price? 
- What body types I can choose? 

At the end of this post, I also provide a simple linear regression model if anyone is interested.

```{r column plots}
# Makes
## Price
tab_make_meanprice <- df_ev[!is.na(price), .(mean_price = round(mean(price), 2)), make][
  order(mean_price)
]
tab_make_medprice <- df_ev[!is.na(price), .(med_price = round(median(price), 2)), make][
  order(med_price)
]
# use reorder to sort
gg_make_meanprice <- tab_make_meanprice %>% 
  ggplot(aes(x = reorder(make, mean_price), y = mean_price, fill = make))+
  geom_col_interactive(aes(tooltip = mean_price, data_id=mean_price))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_y_continuous(labels = comma, breaks = seq(0,300000,25000))+
  labs(x ="Makes", y = "Mean Price", fill = "Makes", title = "Mean Price by Makes")+
  geom_line(aes(x = make, y = 50000, group = 1))+
  geom_line(aes(x = make, y = 100000, group = 1))

gg_make_medprice <- tab_make_medprice %>% 
  ggplot(aes(x = reorder(make, med_price), y = med_price, fill = make))+
  geom_col_interactive(aes(tooltip = med_price, data_id=med_price))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_y_continuous(labels = comma, breaks = seq(0,300000,25000))+
  labs(x ="Makes", y = "Median Price", fill = "Makes", title = "Median Price by Makes")+
  geom_line(aes(x = make, y = 50000, group = 1))+
  geom_line(aes(x = make, y = 100000, group = 1))
```

<br>

## Price by Makes

```{r meanprice, fig.cap='Mean Price by Make',include=TRUE}
ggiraph(ggobj = gg_make_meanprice)
```

```{r medprice,fig.cap='Median Price by Makes',include=TRUE}
ggiraph(ggobj = gg_make_medprice)
```

The median and mean price plots are similar. Renault, MG and Nissan are perfect for a budget below 50,000 AUD. Between 50,000 and 100,000 AUD, we can find some excellent choices with Hyundai, MINI, Kia, BMW, Volvo and Tesla. Beware that Renault only provides Vans in the electric car market in Aus. 

<br> 

```{r}
gg_box_make_price <- df_ev %>% 
  ggplot(aes(x = reorder(make, price), y = price, fill = make))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_y_continuous(labels = comma, breaks = seq(0,400000,25000))+
  geom_line(aes(x = make, y = 50000, group = 1))+
  geom_line(aes(x = make, y = 100000, group = 1))+
  geom_line(aes(x = make, y = 150000, group = 1))+
  labs(x ="Makes", y = "Price", fill = "Makes", title = "Price by Makes")
```

```{r pricemakebox, fig.cap='Price by Makes',include=TRUE, warning=FALSE}
gg_box_make_price
```

Now let us go a bit further to details. Selections for a budget under 50,000 is limited to MG and half of Nissan's offers. However, between 50,000 and 100,000, buyers have a lot more choices including all pure electric vehicles of MINI, Kia, Volvo, Hyundai, Nissan, and approx. 60% EVs of Tesla and BMW. 

On the other hand, buyers should also consider the **EV policies** of states in Aus. For example in **VIC**, the EV policy is: 

"This policy includes an incentive of up to `$3000` to buy an electric vehicle or hydrogen-electric vehicle. The incentive scheme will cover 4000 cars this year and eventually 20,000 EV sales in Victoria over the next three years. This number is unlikely to be surpassed anytime soon, even if the Victorian Government delivers on its policy to up to $10 million worth of EVs." 

The EV market is still minor in Australia while the total number of on-sale EVs on [carsales.com](https://www.carsales.com.au/cars/electric-fueltype/?sort=~Price) is less than 1000. It means, after rough prediction, that every EV buyer will be covered by the policies. 

<br>

## Battery power and travel range

```{r}
df_ev[, .N, .(make, kw)]
unique(df_ev$kw)
gg_point_kw <- df_ev[!is.na(kw)] %>% 
  ggplot(aes(x=make, y=kw, color = make))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  geom_line(aes(x=make, y=100, group =1))+
  geom_line(aes(x=make, y=200, group =1))+
  geom_line(aes(x=make, y=300, group =1))+
  labs(x="Makes",y="Battery KW", color = "Makes", title = "Battery Power (KW) by Makes")
  
gg_point_KW_KM <- df_ev[!is.na(kw) & !is.na(range_km)] %>% 
  ggplot(aes(x=range_km, y=kw, color = make))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5))+
  geom_line(aes(x=range_km, y=100, group =1))+
  geom_line(aes(x=range_km, y=200, group =1))+
  geom_line(aes(x=range_km, y=300, group =1))+
  labs(x="Travel Range (KM)",y="Battery (KW)", color = "Makes", title = "Battery Power(KW) & Travel Range by Makes")
## add
model1 <- df_ev %>% lm(range_km ~ kw, data = .)
summary(model1)

### 
gg_point_range_price <- df_ev[!is.na(price) & !is.na(range_km)] %>% 
  ggplot(aes(x=range_km, y=price, color = make))+
  geom_point_interactive(aes(tooltip = make, data_id=make))+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5))+
  geom_line(aes(y=50000, x=range_km, group =1))+
  geom_line(aes(y=100000, x=range_km, group =1))+
  geom_line(aes(y=150000, x=range_km, group =1))+
  geom_line(aes(x=400, y =price))+
  labs(x="Travel Range (KM)",y="Price (AUD)", color = "Makes", title = "Battery (KW) & Price by Makes")+
  scale_y_continuous(labels = comma, breaks = seq(0, 400000, 50000))
##
model2 <- df_ev %>% lm(price ~ make + range_km + my + odo, data = .)
summary(model2)
```

```{r makekw, fig.cap='Battery Power by Makes', include=TRUE}
gg_point_kw
```

The battery is the core of an EV just like an engine for gasoline cars. 
In the figure \@ref(fig:makekw), The power of the battery generally stays between 100KW and 300KW. 

<br>

```{r kmkw, fig.cap='Battery Power(KW) & Travel Range by Makes', include=TRUE}
gg_point_KW_KM
```

In the figure \@ref(fig:kmkw), the relationship between power and range is quite obvious that the higher battery power the longer travel ranges. A linear regression model demonstrates that **every 1 KW of battery power is estimated to increase travel by 0.6 km with an intercept of 262 km**. 

- range(km) = 262.26380 + 0.63523*kw

<br>

```{r pricerange, fig.cap='Battery Power (KW) by Makes', include=TRUE}
ggiraph(ggobj = gg_point_range_price)
```

In the figure \@ref(fig:pricerange), we can see a clear linear relationship between price and range. Applying make, range, model years and odometers into the linear regression model, we have the summary below: 

```{r model2, include=TRUE}
summary(model2)
```

Most estimations are significant except odometers and model years. It means the odometer and model years are unlikely to have a significant impact on price. This is the biggest difference to gasoline cars. 

The linear regression model, however, demonstrates that **every 1 km increment of the range is estimated to increase 100 AUD on price**. 

<br>

## Body types

```{r}
df_ev[!is.na(price)& !body_type == "", .(mean_price = round(mean(price),2)), .(make, body_type)]
df_ev[body_type == ""] # check empty body type

gg_col_price_body_make <- df_ev[!is.na(price)& !body_type == "", .(mean_price = round(mean(price),2)), .(make, body_type)] %>% 
  ggplot(aes(x=make, y=mean_price, fill = body_type))+
  geom_col_interactive(aes(tooltip = mean_price, data_id = mean_price))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_y_continuous(labels = comma, breaks = seq(0, 600000, 100000))+
  labs(x="Makes",y="Stacked Mean Price (AUD)", fill = "Body Types", title = "Stacked Mean Prices by Body Types and Makes")
  
df_ev[make == "Renault"]
```

```{r pricebodymake, fig.cap='Stacked Mean Prices by Body Types and Makes', include=TRUE}
ggiraph(ggobj = gg_col_price_body_make)
```

In the figure \@ref(fig:pricebodymake),we can find that most of the makes focus on SUVs and Hatch only. Generally, these two body types target family and individuals segments. On the contrary, in the Australian EV market, commercial EVs (larger size vehicles) still have a long ways to go. 